---
title: "Examples for Internal Functions"
author: "NEST CoreDev"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Examples for Internal Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This vignette contains examples for internal classes, methods, and functions of  `teal.slice`. 
For developer use only.

## `FilterStates`

### `init_filter_states`
```{r eval=FALSE}
library(shiny)
df <- data.frame(
  character = letters,
  numeric = seq_along(letters),
  date = seq(Sys.Date(), length.out = length(letters), by = "1 day"),
  datetime = seq(Sys.time(), length.out = length(letters), by = "33.33 hours")
)
rf <- teal.slice:::init_filter_states(
  data = df,
  dataname = "DF"
)
app <- shinyApp(
  ui = fluidPage(
    actionButton("clear", span(icon("xmark"), "Remove all filters")),
    rf$ui_add(id = "add"),
    rf$ui_active("states"),
    verbatimTextOutput("expr"),
  ),
  server = function(input, output, session) {
    rf$srv_add(id = "add")
    rf$srv_active(id = "states")
    output$expr <- renderText({
      deparse1(rf$get_call(), collapse = "\n")
    })
    observeEvent(input$clear, rf$state_list_empty())
  }
)
if (interactive()) {
  shinyApp(app$ui, app$server)
}
```
### `DFFilterStates`
```{r eval=FALSE}
library(shiny)
library(shinyjs)

# create data frame to filter
data_df <- data.frame(
  NUM1 = 1:100,
  NUM2 = round(runif(100, min = 20, max = 23)),
  CHAR1 = sample(LETTERS[1:6], size = 100, replace = TRUE),
  CHAR2 = sample(c("M", "F"), size = 100, replace = TRUE),
  DATE = seq(as.Date("2020-01-01"), by = 1, length.out = 100),
  DATETIME = as.POSIXct(seq(as.Date("2020-01-01"), by = 1, length.out = 100))
)
data_na <- data.frame(
  NUM1 = NA,
  NUM2 = NA,
  CHAR1 = NA,
  CHAR2 = NA,
  DATE = NA,
  DATETIME = NA
)
data_df <- rbind(data_df, data_na)


# initiate `FilterStates` object
filter_states_df <- init_filter_states(
  data = data_df,
  dataname = "dataset",
  datalabel = ("label")
)

ui <- fluidPage(
  useShinyjs(),
  teal.slice:::include_css_files(pattern = "filter-panel"),
  teal.slice:::include_js_files(pattern = "count-bar-labels"),
  column(4, div(
    h4("Active filters"),
    filter_states_df$ui_active("fsdf")
  )),
  column(4, div(
    h4("Manual filter control"),
    filter_states_df$ui_add("add_filters"), br(),
    h4("Condition (i.e. call)"), # display the subsetting call generated by this FilterStates
    textOutput("call_df"), br(),
    h4("Formatted state"), # display human readable filter state
    textOutput("formatted_df"), br()
  )),
  column(4, div(
    h4("Programmatic filter control"),
    actionButton("button1_df", "set NUM1 < 30", width = "100%"), br(),
    actionButton("button2_df", "set NUM2 %in% c(20, 21)", width = "100%"), br(),
    actionButton("button3_df", "set CHAR1 %in% c(\"B\", \"C\", \"D\")", width = "100%"), br(),
    actionButton("button4_df", "set CHAR2 == \"F\"", width = "100%"), br(),
    actionButton("button5_df", "set DATE <= 2020-02-02", width = "100%"), br(),
    actionButton("button6_df", "set DATETIME <= 2020-02-02", width = "100%"), br(),
    hr(),
    actionButton("button7_df", "remove NUM1", width = "100%"), br(),
    actionButton("button8_df", "remove NUM2", width = "100%"), br(),
    actionButton("button9_df", "remove CHAR1", width = "100%"), br(),
    actionButton("button10_df", "remove CHAR2", width = "100%"), br(),
    actionButton("button11_df", "remove DATE", width = "100%"), br(),
    actionButton("button12_df", "remove DATETIME", width = "100%"), br(),
    hr(),
    actionButton("button0_df", "clear all filters", width = "100%"), br()
  ))
)

server <- function(input, output, session) {
  filter_states_df$srv_add("add_filters")
  filter_states_df$srv_active("fsdf")

  output$call_df <- renderPrint(filter_states_df$get_call())
  output$formatted_df <- renderText(filter_states_df$format())

  observeEvent(input$button1_df, {
    filter_state <- teal_slices(teal_slice("dataset", "NUM1", selected = c(0, 30)))
    filter_states_df$set_filter_state(state = filter_state)
  })
  observeEvent(input$button2_df, {
    filter_state <- teal_slices(teal_slice("dataset", "NUM2", selected = c(20, 21)))
    filter_states_df$set_filter_state(state = filter_state)
  })
  observeEvent(input$button3_df, {
    filter_state <- teal_slices(teal_slice("dataset", "CHAR1", selected = c("B", "C", "D")))
    filter_states_df$set_filter_state(state = filter_state)
  })
  observeEvent(input$button4_df, {
    filter_state <- teal_slices(teal_slice("dataset", "CHAR2", selected = c("F")))
    filter_states_df$set_filter_state(state = filter_state)
  })
  observeEvent(input$button5_df, {
    filter_state <- teal_slices(
      teal_slice("dataset", "DATE", selected = c("2020-01-01", "2020-02-02"))
    )
    filter_states_df$set_filter_state(state = filter_state)
  })
  observeEvent(input$button6_df, {
    filter_state <- teal_slices(
      teal_slice("dataset", "DATETIME", selected = as.POSIXct(c("2020-01-01", "2020-02-02")))
    )
    filter_states_df$set_filter_state(state = filter_state)
  })
  observeEvent(input$button7_df, filter_states_df$remove_filter_state(state_id = "NUM1"))
  observeEvent(input$button8_df, filter_states_df$remove_filter_state(state_id = "NUM2"))
  observeEvent(input$button9_df, filter_states_df$remove_filter_state(state_id = "CHAR1"))
  observeEvent(input$button10_df, filter_states_df$remove_filter_state(state_id = "CHAR2"))
  observeEvent(input$button11_df, filter_states_df$remove_filter_state(state_id = "DATE"))
  observeEvent(input$button12_df, filter_states_df$remove_filter_state(state_id = "DATETIME"))
  observeEvent(input$button0_df, filter_states_df$clear_filter_states())
}

if (interactive()) {
  shinyApp(ui, server)
}
```

## `FilterState`

### `init_filter_state`

```{r eval=FALSE}
filter_state <- teal.slice:::init_filter_state(
  x = c(1:10, NA, Inf),
  x_reactive = reactive(c(1:10, NA, Inf)),
  slice = teal_slice(
    varname = "x",
    dataname = "dataname"
  ),
  extract_type = "matrix"
)

shiny::isolate(filter_state$get_call())
app <- shinyApp(
  ui = fluidPage(
    filter_state$ui(id = "app"),
    verbatimTextOutput("call")
  ),
  server = function(input, output, session) {
    filter_state$server("app")

    output$call <- renderText(
      deparse1(filter_state$get_call(), collapse = "\n")
    )
  }
)
if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

### `ChoicesFilterState`

```{r eval=FALSE}
filter_state <- teal.slice:::ChoicesFilterState$new(
  x = c(LETTERS, NA),
  slice = teal_slice(varname = "x", dataname = "data")
)
shiny::isolate(filter_state$get_call())
filter_state$set_state(
  teal_slice(
    dataname = "data",
    varname = "x",
    selected = "A",
    keep_na = TRUE
  )
)
shiny::isolate(filter_state$get_call())
```

```{r eval=FALSE}
# working filter in an app
library(shiny)
library(shinyjs)

data_choices <- c(sample(letters[1:4], 100, replace = TRUE), NA)
attr(data_choices, "label") <- "lowercase letters"
fs <- teal.slice:::ChoicesFilterState$new(
  x = data_choices,
  slice = teal_slice(
    dataname = "data", varname = "variable", selected = c("a", "c"), keep_na = TRUE
  )
)

ui <- fluidPage(
  useShinyjs(),
  teal.slice:::include_css_files(pattern = "filter-panel"),
  teal.slice:::include_js_files(pattern = "count-bar-labels"),
  column(4, div(
    h4("ChoicesFilterState"),
    fs$ui("fs")
  )),
  column(4, div(
    h4("Condition (i.e. call)"), # display the subsetting call generated by this FilterState
    textOutput("condition_choices"), br(),
    h4("Unformatted state"), # display raw filter state
    textOutput("unformatted_choices"), br(),
    h4("Formatted state"), # display human readable filter state
    textOutput("formatted_choices"), br()
  )),
  column(4, div(
    h4("Programmatic filter control"),
    actionButton("button1_choices", "set drop NA", width = "100%"), br(),
    actionButton("button2_choices", "set keep NA", width = "100%"), br(),
    actionButton("button3_choices", "set selection: a, b", width = "100%"), br(),
    actionButton("button4_choices", "deselect all", width = "100%"), br(),
    actionButton("button0_choices", "set initial state", width = "100%"), br()
  ))
)

server <- function(input, output, session) {
  fs$server("fs")
  output$condition_choices <- renderPrint(fs$get_call())
  output$formatted_choices <- renderText(fs$format())
  output$unformatted_choices <- renderPrint(fs$get_state())
  # modify filter state programmatically
  observeEvent(
    input$button1_choices,
    fs$set_state(
      teal_slice(dataname = "data", varname = "variable", keep_na = FALSE)
    )
  )
  observeEvent(
    input$button2_choices,
    fs$set_state(
      teal_slice(dataname = "data", varname = "variable", keep_na = TRUE)
    )
  )
  observeEvent(
    input$button3_choices,
    fs$set_state(
      teal_slice(dataname = "data", varname = "variable", selected = c("a", "b"))
    )
  )
  observeEvent(
    input$button4_choices,
    fs$set_state(
      teal_slice(dataname = "data", varname = "variable", selected = character(0), keep_na = TRUE)
    )
  )
  observeEvent(
    input$button0_choices,
    fs$set_state(
      teal_slice(dataname = "data", varname = "variable", selected = c("a", "c"), keep_na = TRUE)
    )
  )
}

if (interactive()) {
  shinyApp(ui, server)
}
```

### `DateFilterState`

```{r eval=FALSE}
filter_state <- teal.slice:::DateFilterState$new(
  x = c(Sys.Date() + seq(1:10), NA),
  slice = teal_slice(varname = "x", dataname = "data"),
  extract_type = character(0)
)
shiny::isolate(filter_state$get_call())
filter_state$set_state(
  teal_slice(
    dataname = "data",
    varname = "x",
    selected = c(Sys.Date() + 3L, Sys.Date() + 8L),
    keep_na = TRUE
  )
)
shiny::isolate(filter_state$get_call())
```

```{r eval=FALSE}
# working filter in an app
library(shiny)
library(shinyjs)

dates <- c(Sys.Date() - 100, Sys.Date())
data_date <- c(seq(from = dates[1], to = dates[2], length.out = 100), NA)
fs <- teal.slice:::DateFilterState$new(
  x = data_date,
  slice = teal_slice(
    dataname = "data", varname = "x", selected = data_date[c(47, 98)], keep_na = TRUE
  )
)

ui <- fluidPage(
  useShinyjs(),
  teal.slice:::include_css_files(pattern = "filter-panel"),
  teal.slice:::include_js_files(pattern = "count-bar-labels"),
  column(4, div(
    h4("DateFilterState"),
    fs$ui("fs")
  )),
  column(4, div(
    id = "outputs", # div id is needed for toggling the element
    h4("Condition (i.e. call)"), # display the subsetting call generated by this FilterState
    textOutput("condition_date"), br(),
    h4("Unformatted state"), # display raw filter state
    textOutput("unformatted_date"), br(),
    h4("Formatted state"), # display human readable filter state
    textOutput("formatted_date"), br()
  )),
  column(4, div(
    h4("Programmatic filter control"),
    actionButton("button1_date", "set drop NA", width = "100%"), br(),
    actionButton("button2_date", "set keep NA", width = "100%"), br(),
    actionButton("button3_date", "set a range", width = "100%"), br(),
    actionButton("button4_date", "set full range", width = "100%"), br(),
    actionButton("button0_date", "set initial state", width = "100%"), br()
  ))
)

server <- function(input, output, session) {
  fs$server("fs")
  output$condition_date <- renderPrint(fs$get_call())
  output$formatted_date <- renderText(fs$format())
  output$unformatted_date <- renderPrint(fs$get_state())
  # modify filter state programmatically
  observeEvent(
    input$button1_date,
    fs$set_state(teal_slice(dataname = "data", varname = "x", keep_na = FALSE))
  )
  observeEvent(
    input$button2_date,
    fs$set_state(teal_slice(dataname = "data", varname = "x", keep_na = TRUE))
  )
  observeEvent(
    input$button3_date,
    fs$set_state(teal_slice(dataname = "data", varname = "x", selected = data_date[c(34, 56)]))
  )
  observeEvent(
    input$button4_date,
    fs$set_state(teal_slice(dataname = "data", varname = "x", selected = dates))
  )
  observeEvent(
    input$button0_date,
    fs$set_state(
      teal_slice("data", "variable", selected = data_date[c(47, 98)], keep_na = TRUE)
    )
  )
}

if (interactive()) {
  shinyApp(ui, server)
}
```

### `DatetimeFilterState`

```{r eval=FALSE}
filter_state <- teal.slice:::DatetimeFilterState$new(
  x = c(Sys.time() + seq(0, by = 3600, length.out = 10), NA),
  slice = teal_slice(varname = "x", dataname = "data"),
  extract_type = character(0)
)
shiny::isolate(filter_state$get_call())
filter_state$set_state(
  teal_slice(
    dataname = "data",
    varname = "x",
    selected = c(Sys.time() + 3L, Sys.time() + 8L),
    keep_na = TRUE
  )
)
shiny::isolate(filter_state$get_call())
```

```{r eval=FALSE}
# working filter in an app
library(shiny)
library(shinyjs)

datetimes <- as.POSIXct(c("2012-01-01 12:00:00", "2020-01-01 12:00:00"))
data_datetime <- c(seq(from = datetimes[1], to = datetimes[2], length.out = 100), NA)
fs <- teal.slice:::DatetimeFilterState$new(
  x = data_datetime,
  slice = teal_slice(
    varname = "x", dataname = "data", selected = data_datetime[c(47, 98)], keep_na = TRUE
  )
)

ui <- fluidPage(
  useShinyjs(),
  teal.slice:::include_css_files(pattern = "filter-panel"),
  teal.slice:::include_js_files(pattern = "count-bar-labels"),
  column(4, div(
    h4("DatetimeFilterState"),
    fs$ui("fs")
  )),
  column(4, div(
    id = "outputs", # div id is needed for toggling the element
    h4("Condition (i.e. call)"), # display the subsetting call generated by this FilterState
    textOutput("condition_datetime"), br(),
    h4("Unformatted state"), # display raw filter state
    textOutput("unformatted_datetime"), br(),
    h4("Formatted state"), # display human readable filter state
    textOutput("formatted_datetime"), br()
  )),
  column(4, div(
    h4("Programmatic filter control"),
    actionButton("button1_datetime", "set drop NA", width = "100%"), br(),
    actionButton("button2_datetime", "set keep NA", width = "100%"), br(),
    actionButton("button3_datetime", "set a range", width = "100%"), br(),
    actionButton("button4_datetime", "set full range", width = "100%"), br(),
    actionButton("button0_datetime", "set initial state", width = "100%"), br()
  ))
)

server <- function(input, output, session) {
  fs$server("fs")
  output$condition_datetime <- renderPrint(fs$get_call())
  output$formatted_datetime <- renderText(fs$format())
  output$unformatted_datetime <- renderPrint(fs$get_state())
  # modify filter state programmatically
  observeEvent(
    input$button1_datetime,
    fs$set_state(teal_slice(dataname = "data", varname = "x", keep_na = FALSE))
  )
  observeEvent(
    input$button2_datetime,
    fs$set_state(teal_slice(dataname = "data", varname = "x", keep_na = TRUE))
  )
  observeEvent(
    input$button3_datetime,
    fs$set_state(
      teal_slice(dataname = "data", varname = "x", selected = data_datetime[c(34, 56)])
    )
  )
  observeEvent(
    input$button4_datetime,
    fs$set_state(
      teal_slice(dataname = "data", varname = "x", selected = datetimes)
    )
  )
  observeEvent(
    input$button0_datetime,
    fs$set_state(
      teal_slice(
        dataname = "data", varname = "x", selected = data_datetime[c(47, 98)], keep_na = TRUE
      )
    )
  )
}

if (interactive()) {
  shinyApp(ui, server)
}
```

### `EmptyFilterState`

```{r eval=FALSE}
filter_state <- teal.slice:::EmptyFilterState$new(
  x = NA,
  slice = teal_slice(varname = "x", dataname = "data"),
  extract_type = character(0)
)
shiny::isolate(filter_state$get_call())
filter_state$set_state(teal_slice(dataname = "data", varname = "x", keep_na = TRUE))
shiny::isolate(filter_state$get_call())
```

### `FilterStateExpr`

```{r eval=FALSE}
filter_state <- teal.slice:::FilterStateExpr$new(
  slice = teal_slice(
    dataname = "x",
    id = "FA",
    title = "Adult females",
    expr = "sex == 'F' & age >= 18"
  )
)
filter_state$get_call()
```

```{r eval=FALSE}
# working filter in an app
library(shiny)
library(shinyjs)

ui <- fluidPage(
  useShinyjs(),
  teal.slice:::include_css_files(pattern = "filter-panel"),
  teal.slice:::include_js_files(pattern = "count-bar-labels"),
  column(4, div(
    h4("ChoicesFilterState"),
    filter_state$ui("fs")
  )),
  column(8, div(
    h4("Condition (i.e. call)"), # display the subsetting call generated by this FilterState
    textOutput("condition_choices"), br(),
    h4("Unformatted state"), # display raw filter state
    textOutput("unformatted_choices"), br(),
    h4("Formatted state"), # display human readable filter state
    textOutput("formatted_choices"), br()
  ))
)

server <- function(input, output, session) {
  filter_state$server("fs")
  output$condition_choices <- renderPrint(filter_state$get_call())
  output$formatted_choices <- renderText(filter_state$format())
  output$unformatted_choices <- renderPrint(filter_state$get_state())
}

if (interactive()) {
  shinyApp(ui, server)
}
```

### `LogicalFilterState`

```{r eval=FALSE}
filter_state <- teal.slice:::LogicalFilterState$new(
  x = sample(c(TRUE, FALSE, NA), 10, replace = TRUE),
  slice = teal_slice(varname = "x", dataname = "data")
)
shiny::isolate(filter_state$get_call())
filter_state$set_state(
  teal_slice(dataname = "data", varname = "x", selected = TRUE, keep_na = TRUE)
)
shiny::isolate(filter_state$get_call())
```

```{r eval=FALSE}
# working filter in an app
library(shiny)
library(shinyjs)

data_logical <- c(sample(c(TRUE, FALSE), 10, replace = TRUE), NA)
fs <- teal.slice:::LogicalFilterState$new(
  x = data_logical,
  slice = teal_slice(dataname = "data", varname = "x", selected = FALSE, keep_na = TRUE)
)

ui <- fluidPage(
  useShinyjs(),
  teal.slice:::include_css_files(pattern = "filter-panel"),
  teal.slice:::include_js_files(pattern = "count-bar-labels"),
  column(4, div(
    h4("LogicalFilterState"),
    fs$ui("fs")
  )),
  column(4, div(
    id = "outputs", # div id is needed for toggling the element
    h4("Condition (i.e. call)"), # display the subsetting call generated by this FilterState
    textOutput("condition_logical"), br(),
    h4("Unformatted state"), # display raw filter state
    textOutput("unformatted_logical"), br(),
    h4("Formatted state"), # display human readable filter state
    textOutput("formatted_logical"), br()
  )),
  column(4, div(
    h4("Programmatic filter control"),
    actionButton("button1_logical", "set drop NA", width = "100%"), br(),
    actionButton("button2_logical", "set keep NA", width = "100%"), br(),
    actionButton("button3_logical", "set a selection", width = "100%"), br(),
    actionButton("button0_logical", "set initial state", width = "100%"), br()
  ))
)

server <- function(input, output, session) {
  fs$server("fs")
  output$condition_logical <- renderPrint(fs$get_call())
  output$formatted_logical <- renderText(fs$format())
  output$unformatted_logical <- renderPrint(fs$get_state())
  # modify filter state programmatically
  observeEvent(
    input$button1_logical,
    fs$set_state(teal_slice(dataname = "data", varname = "x", keep_na = FALSE))
  )
  observeEvent(
    input$button2_logical,
    fs$set_state(teal_slice(dataname = "data", varname = "x", keep_na = TRUE))
  )
  observeEvent(
    input$button3_logical,
    fs$set_state(teal_slice(dataname = "data", varname = "x", selected = TRUE))
  )
  observeEvent(
    input$button0_logical,
    fs$set_state(
      teal_slice(dataname = "data", varname = "x", selected = FALSE, keep_na = TRUE)
    )
  )
}

if (interactive()) {
  shinyApp(ui, server)
}
```

### `RangeFilterState`

```{r eval=FALSE}
filter_state <- teal.slice:::RangeFilterState$new(
  x = c(NA, Inf, seq(1:10)),
  slice = teal_slice(varname = "x", dataname = "data")
)
shiny::isolate(filter_state$get_call())
filter_state$set_state(
  teal_slice(
    dataname = "data",
    varname = "x",
    selected = c(3L, 8L),
    keep_na = TRUE,
    keep_inf = TRUE
  )
)
shiny::isolate(filter_state$get_call())
```

```{r eval=FALSE}
# working filter in an app
library(shiny)
library(shinyjs)

data_range <- c(runif(100, 0, 1), NA, Inf)
fs <- teal.slice:::RangeFilterState$new(
  x = data_range,
  slice = teal_slice(
    dataname = "data",
    varname = "x",
    selected = c(0.15, 0.93),
    keep_na = TRUE,
    keep_inf = TRUE
  )
)

ui <- fluidPage(
  useShinyjs(),
  teal.slice:::include_css_files(pattern = "filter-panel"),
  teal.slice:::include_js_files(pattern = "count-bar-labels"),
  column(4, div(
    h4("RangeFilterState"),
    fs$ui("fs")
  )),
  column(4, div(
    id = "outputs", # div id is needed for toggling the element
    h4("Condition (i.e. call)"), # display the subsetting call generated by this FilterState
    textOutput("condition_range"), br(),
    h4("Unformatted state"), # display raw filter state
    textOutput("unformatted_range"), br(),
    h4("Formatted state"), # display human readable filter state
    textOutput("formatted_range"), br()
  )),
  column(4, div(
    h4("Programmatic filter control"),
    actionButton("button1_range", "set drop NA", width = "100%"), br(),
    actionButton("button2_range", "set keep NA", width = "100%"), br(),
    actionButton("button3_range", "set drop Inf", width = "100%"), br(),
    actionButton("button4_range", "set keep Inf", width = "100%"), br(),
    actionButton("button5_range", "set a range", width = "100%"), br(),
    actionButton("button6_range", "set full range", width = "100%"), br(),
    actionButton("button0_range", "set initial state", width = "100%"), br()
  ))
)

server <- function(input, output, session) {
  fs$server("fs")
  output$condition_range <- renderPrint(fs$get_call())
  output$formatted_range <- renderText(fs$format())
  output$unformatted_range <- renderPrint(fs$get_state())
  # modify filter state programmatically
  observeEvent(
    input$button1_range,
    fs$set_state(teal_slice(dataname = "data", varname = "x", keep_na = FALSE))
  )
  observeEvent(
    input$button2_range,
    fs$set_state(teal_slice(dataname = "data", varname = "x", keep_na = TRUE))
  )
  observeEvent(
    input$button3_range,
    fs$set_state(teal_slice(dataname = "data", varname = "x", keep_inf = FALSE))
  )
  observeEvent(
    input$button4_range,
    fs$set_state(teal_slice(dataname = "data", varname = "x", keep_inf = TRUE))
  )
  observeEvent(
    input$button5_range,
    fs$set_state(
      teal_slice(dataname = "data", varname = "x", selected = c(0.2, 0.74))
    )
  )
  observeEvent(
    input$button6_range,
    fs$set_state(teal_slice(dataname = "data", varname = "x", selected = c(0, 1)))
  )
  observeEvent(
    input$button0_range,
    fs$set_state(
      teal_slice("data", "variable", selected = c(0.15, 0.93), keep_na = TRUE, keep_inf = TRUE)
    )
  )
}

if (interactive()) {
  shinyApp(ui, server)
}
```


## `FilteredData`

### `FilteredData`

```{r eval=FALSE}
library(shiny)
datasets <- teal.slice:::FilteredData$new(list(iris = iris, mtcars = mtcars))

# get datanames
datasets$datanames()

datasets$set_filter_state(
  teal_slices(teal_slice(dataname = "iris", varname = "Species", selected = "virginica"))
)
isolate(datasets$get_call("iris"))

datasets$set_filter_state(
  teal_slices(teal_slice(dataname = "mtcars", varname = "mpg", selected = c(15, 20)))
)

isolate(datasets$get_filter_state())
isolate(datasets$get_call("iris"))
isolate(datasets$get_call("mtcars"))
### set_filter_state

utils::data(miniACC, package = "MultiAssayExperiment")

datasets <- teal.slice:::FilteredData$new(list(iris = iris, mae = miniACC))

fs <-
  teal_slices(
    teal_slice(
      dataname = "iris", varname = "Sepal.Length", selected = c(5.1, 6.4),
      keep_na = TRUE, keep_inf = FALSE
    ),
    teal_slice(
      dataname = "iris", varname = "Species", selected = c("setosa", "versicolor"),
      keep_na = FALSE
    ),
    teal_slice(
      dataname = "mae", varname = "years_to_birth", selected = c(30, 50),
      keep_na = TRUE, keep_inf = FALSE
    ),
    teal_slice(dataname = "mae", varname = "vital_status", selected = "1", keep_na = FALSE),
    teal_slice(dataname = "mae", varname = "gender", selected = "female", keep_na = TRUE),
    teal_slice(
      dataname = "mae", varname = "ARRAY_TYPE",
      selected = "", keep_na = TRUE, datalabel = "RPPAArray", arg = "subset"
    )
  )
datasets$set_filter_state(state = fs)
shiny::isolate(datasets$get_filter_state())
```


## `FilteredDataset`

### `DataframeFilteredDataset`

```{r eval=FALSE}
library(shiny)
ds <- teal.slice:::DataframeFilteredDataset$new(iris, "iris")
ds$set_filter_state(
  teal_slices(
    teal_slice(dataname = "iris", varname = "Species", selected = "virginica"),
    teal_slice(dataname = "iris", varname = "Petal.Length", selected = c(2.0, 5))
  )
)
isolate(ds$get_filter_state())
isolate(ds$get_call())
```

```{r eval=FALSE}
## set_filter_state
dataset <- teal.slice:::DataframeFilteredDataset$new(iris, "iris")
fs <- teal_slices(
  teal_slice(dataname = "iris", varname = "Species", selected = "virginica"),
  teal_slice(dataname = "iris", varname = "Petal.Length", selected = c(2.0, 5))
)
dataset$set_filter_state(state = fs)
shiny::isolate(dataset$get_filter_state())
```

### `DefaultFilteredDataset`

```{r eval=FALSE}
library(shiny)
ds <- teal.slice:::DefaultFilteredDataset$new(letters, "letters")
isolate(ds$get_filter_state())
isolate(ds$get_call())
```

### `MAEFilteredDataset`

```{r eval=FALSE}
## set_filter_state
utils::data(miniACC, package = "MultiAssayExperiment")
dataset <- teal.slice:::MAEFilteredDataset$new(miniACC, "MAE")
fs <- teal_slices(
  teal_slice(
    dataname = "MAE", varname = "years_to_birth", selected = c(30, 50), keep_na = TRUE
  ),
  teal_slice(
    dataname = "MAE", varname = "vital_status", selected = "1", keep_na = FALSE
  ),
  teal_slice(
    dataname = "MAE", varname = "gender", selected = "female", keep_na = TRUE
  ),
  teal_slice(
    dataname = "MAE", varname = "ARRAY_TYPE", selected = "", keep_na = TRUE
  )
)
dataset$set_filter_state(state = fs)
shiny::isolate(dataset$get_filter_state())
```

#### `calls_combine_by`

```{r eval=FALSE}
calls <- list(
  quote(SEX == "F"), # subsetting on factor
  quote(AGE >= 20 & AGE <= 50), # subsetting on range
  quote(!SURV) # subsetting on logical
)
teal.slice:::calls_combine_by(calls, "&")
```


## Functions

### `countBars`

```{r eval=FALSE}
choices <- sample(as.factor(c("a", "b", "c")), size = 20, replace = TRUE)
counts <- table(choices)
labels <- teal.slice:::countBars(
  inputId = "counts",
  choices = c("a", "b", "c"),
  countsmax = counts,
  countsnow = unname(counts)
)

app <- shinyApp(
  ui = fluidPage(
    div(
      class = "choices_state",
      teal.slice:::include_js_files("count-bar-labels.js"),
      teal.slice:::include_css_files(pattern = "filter-panel"),
      checkboxGroupInput(
        inputId = "choices",
        selected = levels(choices),
        choiceNames = labels,
        choiceValues = levels(choices),
        label = NULL
      )
    )
  ),
  server = function(input, output, session) {
    observeEvent(input$choices, {
      new_counts <- counts
      new_counts[!names(new_counts) %in% input$choices] <- 0
      teal.slice:::updateCountBars(
        inputId = "counts",
        choices = levels(choices),
        countsmax = counts,
        countsnow = unname(new_counts)
      )
    })
  }
)
if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

### `variable_types`

```{r eval=FALSE}
teal.slice:::variable_types(
  data.frame(
    x = 1:3, y = factor(c("a", "b", "a")), z = c("h1", "h2", "h3"),
    stringsAsFactors = FALSE
  ),
  "x"
)

teal.slice:::variable_types(
  data.frame(
    x = 1:3, y = factor(c("a", "b", "a")), z = c("h1", "h2", "h3"),
    stringsAsFactors = FALSE
  ),
  c("x", "z")
)

teal.slice:::variable_types(
  data.frame(
    x = 1:3, y = factor(c("a", "b", "a")), z = c("h1", "h2", "h3"),
    stringsAsFactors = FALSE
  )
)
```

### `get_supported_filter_varnames`

```{r eval=FALSE}
df <- data.frame(
  a = letters[1:3],
  b = 1:3,
  c = Sys.Date() + 1:3,
  d = Sys.time() + 1:3,
  z = complex(3)
)
teal.slice:::get_supported_filter_varnames(df)
```

### `toggle_icon`

```{r eval=FALSE}
library(shiny)

ui <- fluidPage(
  shinyjs::useShinyjs(),
  actionButton("hide_content", label = "hide", icon = icon("xmark")),
  actionButton("show_content", label = "show", icon = icon("check")),
  actionButton("toggle_content", label = "toggle", icon = icon("angle-down")),
  br(),
  div(
    id = "content",
    verbatimTextOutput("printout")
  )
)

server <- function(input, output, session) {
  observeEvent(input$hide_content,
    {
      shinyjs::hide("content")
      teal.slice:::toggle_icon("toggle_content", c("fa-angle-down", "fa-angle-right"), one_way = TRUE)
    },
    ignoreInit = TRUE
  )

  observeEvent(input$show_content,
    {
      shinyjs::show("content")
      teal.slice:::toggle_icon("toggle_content", c("fa-angle-right", "fa-angle-down"), one_way = TRUE)
    },
    ignoreInit = TRUE
  )

  observeEvent(input$toggle_content,
    {
      shinyjs::toggle("content")
      teal.slice:::toggle_icon("toggle_content", c("fa-angle-right", "fa-angle-down"))
    },
    ignoreInit = TRUE
  )

  output$printout <- renderPrint({
    head(faithful, 10)
  })
}
if (interactive()) {
  shinyApp(ui, server)
}
```

### `topological_sort`

```{r eval=FALSE}
teal.slice:::topological_sort(list(A = c(), B = c("A"), C = c("B"), D = c("A")))
teal.slice:::topological_sort(list(D = c("A"), A = c(), B = c("A"), C = c("B")))
teal.slice:::topological_sort(list(D = c("A"), B = c("A"), C = c("B"), A = c()))
```

### `check_ellipsis`

```{r eval=FALSE}
method_class <- function(a, b, c, ...) {
  teal.slice:::check_ellipsis(...)
}
method_class <- function(a, b, c, ...) {
  teal.slice:::check_ellipsis(..., allowed_args = c("y", "z"))
}
```

### `make_c_call`

```{r eval=FALSE}
teal.slice:::make_c_call(1:3)
# [1] 1 2 3

teal.slice:::make_c_call(1)
# [1] 1
```
